Comparing Power Optimizations in the Reorder Buffer

### 2011-06-08 paper/bg.tex
added 1 ref ans an img

\subsection{Distributed ROB}
An accepted optimization for improving the power consumption of an ROB unit 
is to break up the ROB buffer into clusters REF[XXX].  In this organization, the in-flight 
instruction information is distributed across all of the units and then re-assembled 
back into program order after the fact.  The primary advantage of this approach 
is that it dramatically lessens the size of the individual ROB units.  This reduces overall
power by reducing the bit line width as well as unit capacitance.

...IMAGE FROM OUR SLIDES...

@@

The distribution or clustering of the ROB unit also inherently induces a minimal performance overhead 
in particular cases.  By breaking the ROB into smaller pieces and distributing the instruction 
data across the unit, the placement of instructions becomes a performance factor.  Typically, a 
distributed ROB model features only a single write port per ROB cluster.  As a result only one operand 
can be forwarded at a time from the ROB.  In the rare case that an instruction requires two operands be 
forwarded from the same ROB cluster a delay of one cycle is imposed while the operands are forwarded 
sequentially.  It is reported, however, that this happens very rarely in practice and induces less than 1\% 
performance degradation REF[XXX].
\subsection{Latch ROB}
In addition to large power requirement to maintain the entries of the ROB, another 
large power consumer is the read ports of the ROB.  The read port of the ROB is used
to forward the results of an entry to the dispatch unit (DU) to use as an operand.
All though forwarding from the ROB occurs about 6\%\cite{kucuk} of all instructions 
in an N way super scalar processor there are 2N ports (to accommodate FP
 instructions) which add unnecessary complexity and power consumption for a small 
percentage of instructions.  It has been proposed by eliminating the read ports 
in the ROB all together, we can reduce energy consumption\cite{kucuk}.  However
without read ports, a small window appears where the data cannot be accessed,
starting from when the result is placed in to the ROB until the ROB commits that 
entry to the ARF.  To help mitigate stalls that might happen due to this hole, 
we introduce retention latches.   Retention latches are small memory elements and 
we use them to cache the most recent results that have been written into the ROB.
This way instructions are still able to access the most recently produced results
and not stall.  However if the operand needed is not in the ROB, the stream will 
stall until a commit is made to the ARF, and that result value is forwarded to 
the dispatch unit immediately.

>>>>

\subsection{Distributed ROB}
An accepted optimization for improving the power consumption of an ROB unit 
is to break up the ROB buffer into clusters \cite{rabaey}.  In this organization shown in Figure ~\ref{fig:distrob}, 
the in-flight 
instruction information is distributed across all of the units and then re-assembled 
back into program order after the fact.  The primary advantage of this approach 
is that it dramatically lessens the size of the individual ROB units.  This reduces overall
power by reducing the bit line width as well as unit capacitance.

\begin{figure*}
\centering
\includegraphics[height=2in]{clusterROB.png}
\caption{Distributed ROB layout}
\label{fig:distrob}
\end{figure*}

@@

The distribution or clustering of the ROB unit also inherently induces a minimal performance overhead 
in particular cases.  By breaking the ROB into smaller pieces and distributing the instruction 
data across the unit, the placement of instructions becomes a performance factor.  Typically, a 
distributed ROB model features only a single write port per ROB cluster.  As a result only one operand 
can be forwarded at a time from the ROB.  In the rare case that an instruction requires two operands be 
forwarded from the same ROB cluster a delay of one cycle is imposed while the operands are forwarded 
sequentially.  It is reported, however, that this happens very rarely in practice and induces less than 1\% 
performance degradation \cite{rabaey}.
\subsection{Latch ROB}
In addition to large power requirement to maintain the entries of the ROB, another 
large power consumer is the read ports of the ROB.  The read port of the ROB is used
to forward the results of an entry to the dispatch unit (DU) to use as an operand.
All though forwarding from the ROB occurs about 6\%\cite{kucuk} of all instructions 
in an N way super scalar processor there are 2N ports (to accommodate FP
 instructions) which add unnecessary complexity and power consumption for a small 
percentage of instructions.  It has been proposed by eliminating the read ports 
in the ROB all together, we can reduce energy consumption\cite{kucuk}.  However
without read ports, a small window appears where the data cannot be accessed,
starting from when the result is placed in to the ROB until the ROB commits that 
entry to the ARF.  To help mitigate stalls that might happen due to this hole, 
we introduce retention latches.   Retention latches are small memory elements and 
we use them to cache the most recent results that have been written into the ROB.
This way instructions are still able to access the most recently produced results
and not stall.  However if the operand needed is not in the ROB, the stream will 
stall until a commit is made to the ARF, and that result value is forwarded to 
the dispatch unit immediately.
